---
title: "On the eletrodynamic of moving bodies"
author: "André Dogninema Soro, Jean-Romain Roussel, Alexis Achim"
output: 
  pdf_document:
classoption: twocolumn
bibliography: biblio.bibtex
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood density is among the key traits as it is related to mechanical strength and dimensional stability. Variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In the present study, we aim at disentangling age, genetic and environmental influences on the pith-to-bark variation of wood density. 
We analyzed the wood density profiles of a 16-year-old genetic experimentation plantation from white spruce seedlings grown in nurseries and planted at two years of age. repeated in two contrasting environments (Asselin and Saint-Casimir) in the province of Quebec. A total of 2365 trees belonging to 93 full-sib families were sampled for this study. We developed a model to describe the interannual variation of the wood density from  pith-to-bark. Then, we developed a variability index from the model residues to determine the variation in the density of pith-to-bark in order to identify tree families that are least sensitive to variations in climatic conditions. The genetic control of density, ring width,  model parameters and uniformity index was evaluated. This makes it possible to see the trait the level of heritability of each of these characters and thus to know if it is useful to select trees from orbservations made on these trees. ASReml3 package in R was used to estimate trait genetic parameters based on the residual maximum likelihood. Strong heritability has been observed in density, ring width and model parameters have heritability ranging from low to moderate, and the index of variability has almost no heritability. Future work can be done on older trees and also on one-time events to see if this changes the heritability of the variability index.\

\clearpage

# Introduction

Thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@X, @Y]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@X, @Y]. \ 

Improvement programs by genetic selection are in place for the most important commercial species [@X,@Y] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@X,@Y]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions (Canadian Council of Forest Ministers 2009). \ 

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Koubaa2000; @X,@Y] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Consequently studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Li1997]. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. Several traits were considered for wood quality assessment and the density is one of the most widely used indicators and arguably the most important one [@X, @Y] because this trait is important for both components of quality. \ 

With regard to the biological component, in has been demonstrated [@X, @Y] that trees with a more uniform wood densitiy within the stem are also the more resistant to mortality. For example a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bioenergy [@Macdonald2002; @Mckendry2002]. For exemple, a strong interannual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance. It is therefore important from both a biological point of view and a end-use point of view to have a wood with a uniform density within the stem. \ 

From these studies we understand that there might be a strong gain at selecting trees that are prone to produce wood of uniform density to get a resistant and higly valuable forest in the next decades. To reach this goal is is important to understant where do the variations of density come from. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011]. First a decrease of density during the first years of growth and then an increase from the beginning of the formation of mature wood [@panshin1980]. In addition this radial variations there are hight frequency interannual variation [@Nicault2010]. Contrary to the radial variation that is under internal control, the interannual variation of the density is influenced mainly by events events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. \ 

Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b, @Beaulieu2006] and this makes this trait attractive for genetic improvement. 

<!-- Ca dit tout l'inverse de ce que veux dire... je sais quoi faire de ca. -->
<!-- It should be noted that although density is predominantly under genetic control, it is partly controlled by the environment as growth is density dependent [@Pape1999] and growth is predominantly environmentally controlled. \ -->

However, the timeframe of such improved forest management being long it is essential to take into account the current climate change context [@X,@Y]. With the increasing temperature, the trees are expected to growth faster and thus being of lower quality [@X,@Y]. In addition to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accuralty, with climate change. \ 

In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and clone individus of interest that are both insensitive to climate change and genetically prone to make homogenous wood.

<!-- objectif pas clair: toujours la même question: pourquoi ? -->
<!-- In  this study, we tested using a model similar to that of @Xiang2014 to subtract the effect of cambial age from wood density in order to assess the interannual variation related to environmental conditions. The specific objectives were: (1) present the development of a model which describes the general white spruce density annual ring number from the pith, (2) characterizing the genetic influence on wood density variation in white spruce. In this way, we will be able to identify tree families that are least sensitive to interannual variations in climate. -->

\clearpage

# Material and method

\clearpage

# Conclusion
This study assessed the variability of wood density in white spruce. Using the optimal function of the stats package to find the best parameters iteratively for each tree allowed to have an adjustment for the density profile of each of the trees in the study. Thus, the index of variability of the density was determined from the profile of each tree. The low heritability of this index puts a damper on its use for identifying potential tree family of interest for breeding programs from a low interannual variability perspective of wood density. As for the heritability of ring density and width, it is similar to what can be found in previous studies. The application of the method used in this study to older trees can be interesting in order to make a comparison and see if the heritability is the same and if this index can be used in this case to select families of trees for genetic improvement.

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

\clearpage

# References
